Using Metric Space Indexing for Complete and Efficient Record Linkage

### 2017-11-22 paper.tex
Update on Overleaf.

Record linkage is the process of identifying records that refer to the
same real-world entities, in situations where entity identifiers are
unavailable. Records are linked on the basis of similarity between
common attributes, with every pair being classified as a link or
non-link depending on their degree of similarity. Record linkage is
usually performed in a three-step process: first groups of similar
candidate records are identified using indexing; pairs within the same
group are then compared in more detail, and finally classified.
%
Even state-of-the-art indexing techniques, such as Locality Sensitive
Hashing, have potential drawbacks. They may fail to group together some
true matching records with high similarity. Conversely, they may group
records with low similarity, leading to high computational overhead.
%
We propose using metric space indexing to perform \emph{complete} record
linkage, which results in a parameter-free record linkage process
combining indexing, comparison and classification into a single step
delivering complete and efficient record linkage. Our experimental
evaluation on real-world datasets from several domains shows that
linkage using metric space indexing can yield better quality than
current indexing techniques, with similar execution cost, without the
need for domain knowledge or trial and error to configure the process.

@@

Record linkage, also known as entity resolution, data matching and
duplicate detection~\cite{Chr12}, is the process of identifying and
matching records that refer to the same real-world entities within or
across datasets. The entities to be linked are often people (such as
patients in hospital or customers in business datasets), but record
linkage can also be applied to link consumer products or bibliographic
records~\cite{Chr12}. Record linkage is commonly challenged by the lack
of unique entity identifiers (keys) in the datasets to be linked, which
prevents the use of a database join. Instead, the linkage of records
requires the comparison of the common attributes (or fields) that are
available within the datasets. For datasets that contain information
about individuals, these attributes include the names, address, dates of
birth, and so on, of individuals.

To overcome data quality issues such as typographical errors and
variations (which are common in name and address values~\cite{Chr12}),
approximate string comparison functions (such as edit distance, the
Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used to
compare pairs of records, leading to a vector of similarities (one
similarity per attribute/field compared) for each pair. These similarity
vectors are then used to classify the compared record pairs into
\emph{links} (where it is assumed both records in a pair correspond to
the same real-world entity) and \emph{non-links} (where the records are
assumed to correspond to two different entities). Various
classification methods have been employed in record
linkage~\cite{Chr12,Don15}, ranging from simple threshold-based to
sophisticated clustering, supervised classification techniques, and
active learning approaches~\cite{Wan15}.

Besides the issues of a lack of unique entity identifiers, and data
quality (which will affect linkage quality), record linkage is also
challenged by the increasing sizes of the datasets to be
linked~\cite{Don15}. To avoid full pair-wise comparison of all possible
record pairs (quadratic in the sizes of the datasets to be linked),
blocking techniques, commonly known as \emph{indexing}~\cite{Chr12b},
are used. These split the datasets into smaller blocks in a
computationally efficient way, such that records that are likely to
correspond to the same entity are grouped into the same block. Only
records within the same block are then compared in more detail.

@@

Techniques to link records across datasets have been investigated for
over five decades~\cite{Fel69,New59}, with the scalability of linking
being an ongoing challenge as datasets grow in size and complexity.
Traditional blocking~\cite{Chr12b} uses a set of attributes (known as a
\emph{blocking key}) to insert records that share the same value(s) in
their blocking key into the same block. Only records within the same
block are  compared to each other. To overcome variations and
misspellings, the values used in blocking keys can be phonetically
encoded using functions such as Soundex, NYSIIS, or
Double-Metaphone~\cite{Chr12}. These convert a string into a code
according to its pronunciation, assigning the same code to similar
sounding names (such as `Gail' and `Gayle'). Multiple blocking keys
may be used to deal with the problem of missing attribute values.

@@

Metric space indexing (MSI) techniques provide indexing structures to
support the comparison of records in one set with those in another,
typically also offering similarity search operations. To create an MSI
data structure, it is necessary to define a distance measure between
records, with certain properties including the \emph{triangle
inequality}. Similarity search operations include
\emph{range-search(\textbf{q},d)}, where all records $\mathbf{r}$
in a database $\mathbf{D}$ within a distance $d$ of a query record
$\mathbf{q}$ are identified; \emph{nearest-neighbour(\textbf{q})},
returning the record $\mathbf{r} \in \mathbf{D}$ with smallest
distance to $\mathbf{q}$; and \emph{nearest-n(\textbf{q},n)}, returning
the $n$ closest records in $\mathbf{D}$ to $\mathbf{q}$.
\cite{Zezula2010} gives an overview of MSI techniques. Here we choose
one metric space indexing structure, the M-tree~\cite{Ciaccia97indexingmetric},
and investigate its efficacy for record linkage.

@@

\smallskip
\textbf{Brute force:}
Two nested loops are used to compare every record in
$\mathbf{D}_A$ with every record in $\mathbf{D}_B$. Each pair is
classified as a link if the distance $d$ between the records is less
than the threshold. This approach is guaranteed to identify all links,
with complexity $ O(|\mathbf{D}_A| \times |\mathbf{D}_B|) $.

@@

\smallskip
\textbf{M-tree:} Besides the distance threshold, $d$, the M-tree
linkage algorithm has no additional parameters. In a similar manner
to the \emph{LSH-MinHash} approach, each record in $\mathbf{D}_A$ is inserted into an M-tree. To perform linkage, the algorithm iterates
over each record in $\mathbf{b} \in \mathbf{D}_B$. A
\emph{range-search(\textbf{b},d)} operation is performed, passing the
distance threshold as the second parameter. All the returned records
are directly classified as links.

@@

%\smallskip
%\textbf{Datasets:}
\label{sec-data}
We used three datasets from two domains in our experiments, as
summarised in Table~\ref{table-datasets}. The first is
\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer
to 112 machine learning publications. Cora is commonly used as a
benchmark dataset in the literature for assessing linkage algorithms.
Ground truth is provided via a unique \emph{paper\_id} identifier of
the form ``blum1993". In this experiment linkage is performed over
the same set of records.

The other two datasets are historical Scottish records of vital events
(birth, marriages and deaths) one registered on the \emph{Isle of
Skye}, a rural district, and the other records from
\emph{Kilmarnock}, an industrial town. These datasets were created by
historical demographers who extensively curated and linked both
datasets~\cite{reid2002,reid2006}. Both datasets include the names,
gender, addresses of individuals and their parents. Ground truth was
generated by the demographers based on their extensive domain
knowledge.

In all of our experiments we use a single distance metric which is the
sum of the Levenshtein~\cite{Levenshtein66} distances between the set
of attributes being compared.

@@

In this section, we perform linkage on the Cora dataset using all four
approaches presented in this paper: brute force, traditional blocking,
LSH and M-tree, using several selected configurations for the blocking
and LSH approaches. The distance threshold is varied between $0 \le d \le 250$~\footnote{Relatively high Levenshtein distances are included
since Cora includes a number of low-similarity true matches.}. For
traditional blocking, the following Cora attributes are used
individually as blocking keys: \emph{author}, \emph{title},
\emph{venue}, \emph{location}, \emph{publisher} and \emph{year}. We
also use a combined blocking key comprising all attributes.

\Cref{cora-quality} shows the precision, recall, and F-measure~\cite{Chr12} for various thresholds. As expected, low 
thresholds give high precision and low recall, and the reverse for high
threshold values. Brute force and M-tree give identical results, as
expected. The best linkage quality, with an F-measure of around 0.7,
is achieved by a number of linkers, including brute force, M-tree, blocking on \emph{authors} and blocking on all attributes, and two
of the LSH configurations. All of these give similar overall results,
apart from blocking on \emph{authors}, which gives much better quality
at very high distance threshold values. This is due to the incomplete
nature of the approach, avoiding comparisons of significant numbers
of high-similarity non-matches and thus avoiding these becoming false
positives and keeping precision high.

@@

\Cref{comparison-of-results-cora} shows greater detail for selected
linkers, showing the parameters for the experiment, the number of
distance comparisons made, the precision, recall and the F-measure
achieved by each algorithm. In the \emph{Linker} column the algorithm
name is followed by its parameters: for LSH the number of the bands
followed by the band size, and for traditional blocking the attributes
used for blocking. The number of distance comparisons is reported as a
machine-independent proxy for execution cost, since code profiling
shows that distance calculations are dominant.

@@

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
\vspace{-6mm}
\caption{Blocking on `authors' attr.}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
\vspace{-6mm}
\caption{Blocking on `title' attr.}
\end{subfigure} \vspace{5mm}

@@

\caption{Linkage results on the Cora dataset}
\label{cora-quality}
\end{figure}

@@

\begin{table}[ht]
\caption{Linkage quality on Cora dataset with distance threshold
  $d= 70$.}
  \label{comparison-of-results-cora}
  \centering
  \begin{footnotesize}
  \begin{tabular}{lrrrr} \hline\noalign{\smallskip}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68  \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree              &  902,693                  & 0.84      & 0.57   & 0.68       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
LSH-5-2            &  342,849	& 0.91	& 0.55	& 0.69       \\
LSH-10-2           &  513,947	& 0.88	& 0.57	& 0.69       \\
LSH-2-5            &  14,329	& 0.99	& 0.28	& 0.43      \\
LSH-5-5            &  22,057	& 0.99	& 0.36	& 0.53       \\
LSH-10-5           &  26,167	& 0.98	& 0.4	& 0.57      \\
LSH-2-10           &  4,711	& 1	& 0.15	& 0.27    \\
LSH-5-10           &  6,501	& 1	& 0.19	& 0.32      \\
LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
Block-year         &  115,893                  & 0.99      & 0.35   & 0.51       \\
Block-authors      &   11,039                   & 0.94      & 0.16   & 0.28       \\
Block-title        &   27,407                   & 0.95      & 0.42   & 0.58       \\
Block-venue        &   36,647                   & 0.85      & 0.29   & 0.44       \\
Block-location     & 1,009,957                 & 0.83      & 0.43   & 0.57       \\
Block-publisher    &  833,079                  & 0.85      & 0.44   & 0.58       \\
Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

\Cref{comparison-of-results-demography-skye} and
\Cref{comparison-of-results-demography-kili} show greater detail for
selected linkers. In both cases the F-measure achieved is better for
M-tree than any of the LSH linkers. The better linkage quality achieved
by M-tree is largely due to recall for M-tree being much higher than for
any of the LSH configurations. In most cases, LSH out-performs M-tree in
terms of precision. More significantly, LSH linkage quality is heavily
dependent on the configuration parameters. For plausible settings for
the number of bands and band size, F-measure varies from $0.01$
(extremely poor) to $0.47$ (relatively good) for Skye and from $0.03$
to $0.49$ for Kilmarnock. In both cases \emph{LSH-10-2} performs best,
but since this is data-dependent there is no guarantee that these
parameters would work well with another dataset.

The number of distance comparisons varies dramatically among the various
linkers. M-tree always performs the most comparisons, since they are
intrinsic to the \emph{range-search} algorithm. The core part of the LSH
linker performs Jaccard similarity comparisons and hashing; distance
comparisons are only performed in the final step to determine whether a
candidate pair is a link. The LSH configurations yielding the best
results perform distance comparisons of the same order of magnitude as
M-tree. This indicates that the good LSH linkers return many candidates
beyond the distance threshold. Thus, in order to get good results, LSH
tends towards a brute force search over the candidate results. Despite
this, LSH is much faster due to the efficiency of the hashing process.

@@

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Skye dataset\label{skye-quality-mtree}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Kilmarnock dataset\label{kilmarnock-quality-mtree}}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
\vspace{-6mm}
\caption{F1 Measure on Skye dataset \\ with M-tree and all LSH configurations}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
\vspace{-6mm}
\caption{F1 Measure on Kilmarnock dataset \\ with M-tree and all LSH configurations}
\end{subfigure}
\caption{Linkage results on the demographic datasets}
\label{demography-quality}
\end{figure}

@@

\begin{table}[ht]
\caption{Linkage quality on Skye dataset with distance threshold
$d= 2$.}
\label{comparison-of-results-demography-skye}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrr} \hline\noalign{\smallskip}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\ \hline
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2  & 3,109,250                   & 0.63      & 0.03   & 0.06       \\

@@

\begin{table}[ht]
\caption{Linkage quality on Kilmarnock dataset with distance
threshold $d= 2$.}
\label{comparison-of-results-demography-kili}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrrrr} \hline\noalign{\smallskip}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338                   & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848                   & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368                      & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ 
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

We note that the good results yielded by both traditional blocking and
LSH are partly due to the fact that (in the limit) they tend towards
brute force as the number of records in the block increase. A second,
unexpected, result is that illustrated in
\Cref{comparison-of-results-cora}, namely that incomplete approaches
such as LSH can in some cases yield higher precision than that achieved
by a complete method such as M-tree. This is due to the incomplete
linker's ability to mask the inability of a classifier based solely on
record similarity to correctly classify high-similarity non-matches or
low-similarity matches.

The observation that LSH can yield results quickly at the cost of some
recall suggests that hybrid algorithms may be a potentially profitable
future direction for exploration. We hypothesise that it may be possible
to use an incomplete technique like LSH to form indices into a data
structure such as an M-tree. In this manner, M-tree nodes could be used
as the starting point for range-search queries and, in so doing, prune
the search space considerably to yield high quality results quickly.

>>>>

Record linkage is the process of identifying records that refer to the
same real-world entities, in situations where entity identifiers are
unavailable. Records are linked on the basis of similarity between
common attributes, with every pair being classified as a link or
non-link depending on their degree of similarity. Record linkage is
usually performed in a three-step process: first groups of similar
candidate records are identified using indexing, pairs within the same
group are then compared in more detail, and finally classified.
%
Even state-of-the-art indexing techniques, such as Locality Sensitive
Hashing, have potential drawbacks. They may fail to group together some
true matching records with high similarity. Conversely, they may group
records with low similarity, leading to high computational overhead.
%
We propose using metric space indexing to perform \emph{complete} record
linkage, which results in a parameter-free record linkage process
combining indexing, comparison and classification into a single step
delivering complete and efficient record linkage. Our experimental
evaluation on real-world datasets from several domains shows that
linkage using metric space indexing can yield better quality than
current indexing techniques, with similar execution cost, without the
need for domain knowledge or trial and error to configure the process.

@@

Record linkage, also known as entity resolution, data matching and
duplicate detection~\cite{Chr12}, is the process of identifying and
matching records that refer to the same real-world entities within or
across datasets. The entities to be linked are often people (such as
patients in hospital or customers in business datasets), but record
linkage can also be applied to link consumer products or bibliographic
records~\cite{Chr12}. Record linkage is commonly challenged by the lack
of unique entity identifiers (keys) in the datasets to be linked, which
prevents the use of a database join. Instead, the linkage of records
requires the comparison of the common attributes (or fields) that are
available within the datasets. For datasets that contain information
about individuals, these attributes include the names, addresses,
dates of birth, and so on, of individuals.

To overcome data quality issues such as typographical errors and
variations (which are common in name and address values~\cite{Chr12}),
approximate string comparison functions (such as edit distance, the
Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used to
compare pairs of records, leading to a vector of similarities (one
similarity per attribute compared) for each pair. These similarity
vectors are then used to classify the compared record pairs into
\emph{links} (where it is assumed both records in a pair correspond to
the same real-world entity) and \emph{non-links} (where the records are
assumed to correspond to two different entities). Various
classification methods have been employed in record
linkage~\cite{Chr12,Don15}, ranging from simple threshold-based to
sophisticated clustering, supervised classification techniques, and
active learning approaches~\cite{Wan15}.

Besides the issues of a lack of unique entity identifiers and data
quality (which will affect linkage quality), record linkage is also
challenged by the increasing sizes of the datasets to be
linked~\cite{Don15}. To avoid full pair-wise comparison of all possible
record pairs (quadratic in the sizes of the datasets to be linked),
blocking techniques, commonly known as \emph{indexing}~\cite{Chr12b},
are used. These split the datasets into smaller blocks in a
computationally efficient way, such that records that are likely to
correspond to the same entity are grouped into the same block. Only
records within the same block are then compared in more detail.

@@

Techniques to link records across datasets have been investigated for
over five decades~\cite{Fel69,New59}, with the scalability of linking
being an ongoing challenge as datasets grow in size and complexity.
Traditional blocking~\cite{Chr12b} uses a set of attributes (known as a
\emph{blocking key}) to insert records that share the same value(s) in
their blocking key into the same block. Only records within the same
block are  compared to each other. To overcome variations and
misspellings, the values used in blocking keys can be phonetically
encoded using functions such as Soundex, NYSIIS, or
Double-Metaphone~\cite{Chr12}. These convert a string into a code
according to its pronunciation, assigning the same code to similar
sounding names (such as `Gail' and `Gayle'). Multiple blocking keys
may also be used to deal with the problem of missing attribute values.

@@

Metric space indexing (MSI) techniques~\cite{Zezula2010} provide
indexing structures to support the comparison of records in one set
with those in another, typically also offering similarity search
operations. To create an MSI data structure, it is necessary to define
a distance measure between records, with certain properties including
the \emph{triangle inequality}~\cite{Zezula2010}. Similarity search
operations include \emph{range-search(\textbf{q},d)}, where all
records $\mathbf{r}$ in a database $\mathbf{D}$ within a distance $d$
of a query record $\mathbf{q}$ are identified;
\emph{nearest-neighbour(\textbf{q})}, returning the record $\mathbf{r}
\in \mathbf{D}$ with smallest distance to $\mathbf{q}$; and
\emph{nearest-n(\textbf{q},n)}, returning the $n$ closest records in
$\mathbf{D}$ to $\mathbf{q}$.
%\cite{Zezula2010} gives an overview of MSI techniques.
Here we choose one metric space indexing structure, the
M-tree~\cite{Ciaccia97indexingmetric}, and investigate its efficacy
for record linkage.

@@

\smallskip
\textbf{Brute force:}
Two nested loops are used to compare every record in
$\mathbf{D}_A$ with every record in $\mathbf{D}_B$. Each pair is
classified as a link if the distance between the records is equal to
or less than the threshold $d$. This approach is guaranteed to identify
all links, with complexity $ O(|\mathbf{D}_A| \times |\mathbf{D}_B|)$.

@@

\smallskip
\textbf{M-tree:} Besides the distance threshold, $d$, the M-tree
linkage algorithm has no additional parameters. In a similar manner
to the \emph{LSH-MinHash} approach, each record in $\mathbf{D}_A$ is inserted into an M-tree. To perform linkage, the algorithm iterates
over each record $\mathbf{b} \in \mathbf{D}_B$. A
\emph{range-search(\textbf{b},d)} operation is performed on the
M-tree, passing the distance threshold $d$ as the second parameter.
All the returned records are directly classified as links.

@@

%\smallskip
%\textbf{Datasets:}
\label{sec-data}
We used three datasets from two domains in our experiments, as
summarised in Table~\ref{table-datasets}. The first is
\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer
to 112 machine learning publications. Cora is commonly used as a
benchmark dataset in the literature for assessing linkage algorithms.
Ground truth is provided via a unique \emph{paper\_id} identifier of
the form ``blum1993". In this experiment linkage is performed over
the same set of records (i.e.\ a de-duplication~\cite{Chr12}).

The other two datasets are historical Scottish records of vital events
(birth, marriages and deaths) one registered on the \emph{Isle of
Skye}, a rural district, and the other records from
\emph{Kilmarnock}, an industrial town. These datasets were created by
historical demographers who extensively curated and linked both
datasets~\cite{reid2006,reid2002}. Both datasets include the names,
gender, addresses of individuals and their parents. Ground truth was
generated by the demographers based on their extensive domain
knowledge.

In all of our experiments we use a single distance metric which is the
sum of the Levenshtein~\cite{Levenshtein66} edit distances between the
set of attributes being compared.

@@

In this section, we perform linkage on the Cora dataset using all four
approaches presented in this paper: brute force, traditional blocking,
LSH and M-tree, using several selected configurations for the blocking
and LSH approaches. The distance threshold is varied between $0 \le d
\le 250$~\footnote{Relatively high Levenshtein edit distances are
included since Cora includes a number of low-similarity true
matches.}. For traditional blocking, the following Cora attributes are
used individually as blocking keys: \emph{author}, \emph{title},
\emph{venue}, \emph{location}, \emph{publisher} and \emph{year}. We
also use a combined blocking key comprising all attributes.

\Cref{cora-quality} shows the precision, recall, and F-measure~\cite{Chr12} for various thresholds. As expected, low 
thresholds give high precision and low recall, and the reverse for high
threshold values. Brute force and M-tree give identical results, as
expected. The best linkage quality, with an F-measure of around 0.7,
is achieved by a number of linkers, including brute force, M-tree, blocking on \emph{authors} and blocking on all attributes, and two
of the LSH configurations. All of these give similar overall results,
apart from blocking on \emph{authors}, which gives much better quality
at very high distance thresholds. This is due to the incomplete
nature of the approach, avoiding comparisons of significant numbers
of high-similarity non-matches and thus avoiding these becoming false
positives and keeping precision high.

@@

\Cref{comparison-of-results-cora} shows greater detail for selected
linkers, showing the parameters for the experiment, the number of
distance comparisons made, and the precision, recall and F-measure
achieved by each algorithm. In the \emph{Linker} column the algorithm
name is followed by its parameters: for LSH the number of the bands
followed by the band size, and for traditional blocking the attributes
used for blocking. The number of distance comparisons is reported as a
machine-independent proxy for execution cost, since code profiling
shows that distance calculations are dominant.

@@

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
\vspace{-6mm}
\caption{Blocking on `authors' attribute}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
\vspace{-6mm}
\caption{Blocking on `title' attribute}
\end{subfigure} \vspace{5mm}

@@

\caption{Linkage results on the Cora dataset.}
\label{cora-quality}
\end{figure}

@@

\begin{table}[ht]
\caption{Linkage quality on Cora dataset with distance threshold
  $d= 70$.}
  \label{comparison-of-results-cora}
  \centering
  \begin{footnotesize}
  \begin{tabular}{lrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
%\parbox{0.2\textwidth}{\centering Linker} &
%\parbox{0.2\textwidth}{\centering ~~~Comparisons} &
%\parbox{0.17\textwidth}{\centering ~~~Precision} &
%\parbox{0.17\textwidth}{\centering ~~~Recall} &
%\parbox{0.17\textwidth}{\centering ~~~F1 Measure} \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68  \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree              &  902,693                  & 0.84      & 0.57   & 0.68       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
LSH-5-2            &  342,849	& 0.91	& 0.55	& 0.69       \\
LSH-10-2           &  513,947	& 0.88	& 0.57	& 0.69       \\
LSH-2-5            &  14,329	& 0.99	& 0.28	& 0.43      \\
LSH-5-5            &  22,057	& 0.99	& 0.36	& 0.53       \\
LSH-10-5           &  26,167	& 0.98	& 0.4	& 0.57      \\
LSH-2-10           &  4,711	& 1	& 0.15	& 0.27    \\
LSH-5-10           &  6,501	& 1	& 0.19	& 0.32      \\
LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
Block-year         &  115,893                  & 0.99      & 0.35   & 0.51       \\
Block-authors      &   11,039                   & 0.94      & 0.16   & 0.28       \\
Block-title        &   27,407                   & 0.95      & 0.42   & 0.58       \\
Block-venue        &   36,647                   & 0.85      & 0.29   & 0.44       \\
Block-location     & 1,009,957                 & 0.83      & 0.43   & 0.57       \\
Block-publisher    &  833,079                  & 0.85      & 0.44   & 0.58       \\
Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

Tables~\ref{comparison-of-results-demography-skye}
and~\ref{comparison-of-results-demography-kili} show greater detail for
selected linkers. In both cases the F-measure achieved is better for
M-tree than any of the LSH linkers. The better linkage quality achieved
by M-tree is largely due to recall for M-tree being much higher than for
any of the LSH configurations. In most cases, LSH out-performs M-tree in
terms of precision. More significantly, LSH linkage quality is heavily
dependent on the configuration parameters. For plausible settings for
the number of bands and band size, F-measure varies from $0.01$
(extremely poor) to $0.47$ (relatively good) for Skye and from $0.03$
to $0.49$ for Kilmarnock. In both cases \emph{LSH-10-2} performs best,
but since this is data-dependent there is no guarantee that these
parameters would work well with another dataset.

The number of distance comparisons varies dramatically among the various
linkers. M-tree always performs the most comparisons, since they are
intrinsic to the \emph{range-search} algorithm. The core part of the LSH
linker performs Jaccard similarity comparisons and hashing; distance
comparisons are only performed in the final step to determine whether a
candidate pair is a link. The LSH configurations yielding the best
results perform distance comparisons of the same order of magnitude as
M-tree. This indicates that the good LSH linkers return many candidates
beyond the distance threshold. Thus, in order to get good results, LSH
tends towards a brute force search over the candidate results. Despite
this, LSH is faster due to the efficiency of the hashing process.

@@

\begin{figure}[t]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Skye dataset\label{skye-quality-mtree}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Kilmarnock dataset\label{kilmarnock-quality-mtree}}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
\vspace{-6mm}
\caption{F1 Measure on Skye dataset \\ with M-tree and all LSH configurations}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
\vspace{-6mm}
\caption{F1 Measure on Kilmarnock dataset \\ with M-tree and all LSH configurations}
\end{subfigure}
\caption{Linkage results on the demographic datasets.}
\label{demography-quality}
\end{figure}

@@

\begin{table}[t]
\caption{Linkage quality on Skye dataset with distance threshold
$d= 2$.}
\label{comparison-of-results-demography-skye}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
%\parbox{0.2\textwidth}{\centering Linker} &
%\parbox{0.2\textwidth}{\centering Comparisons} &
%\parbox{0.17\textwidth}{\centering Precision} &
%\parbox{0.17\textwidth}{\centering Recall} &
%\parbox{0.17\textwidth}{\centering F1 Measure} \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2  & 3,109,250                   & 0.63      & 0.03   & 0.06       \\

@@

\begin{table}[t]
\caption{Linkage quality on Kilmarnock dataset with distance
threshold $d= 2$.}
\label{comparison-of-results-demography-kili}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrrrr} \hline\noalign{\smallskip}
  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
%\parbox{0.2\textwidth}{\centering Linker} &
%\parbox{0.2\textwidth}{\centering Comparisons} &
%\parbox{0.17\textwidth}{\centering Precision} &
%\parbox{0.17\textwidth}{\centering Recall} &
%\parbox{0.17\textwidth}{\centering F1 Measure} \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338                   & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848                   & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368                      & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ 
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

We note that the good results yielded by both traditional blocking and
LSH are partly due to the fact that (in the limit) they tend towards
brute force as the number of records in the block increase. A second,
unexpected, result is that illustrated in
\Cref{comparison-of-results-cora}, namely that incomplete approaches
such as LSH can in some cases yield higher precision than that achieved
by a complete method such as M-tree. This is due to the incomplete
linker masking the inability of a classifier based solely on
record similarity to correctly classify high-similarity non-matches or
low-similarity matches.

The observation that LSH can yield results quickly at the cost of some
recall suggests that hybrid algorithms may be a potentially profitable
future direction for exploration. We hypothesise that it may be possible
to use an incomplete technique like LSH to form indices into a data
structure such as an M-tree. In this manner, M-tree nodes could be used
as the starting point for range-search queries and, in so doing, prune
the search space considerably to yield high quality results quickly.
We plan to develop such hybrid indexing approaches with the aim to
achieve fully automated, complete and efficient record linkage.
